---
title: "MOCC Analysis"
author: "Guoqing Diao"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  word_document: default
  html_document: default
---

```{r pressure, echo=FALSE}
library(MASS)
library(sandwich)
library(ggplot2)
```

```{r, echo=TRUE}
setwd("/Users/gdiao/Dropbox/GWU/BSC/NIH/MOCC/Data")

### Read data provided by Alex 
mocc  <- arrow::read_feather("mocc_update_1-10-25.feather")

### Check variable names
colnames(mocc)

### Frequencies of surge decile 
table(mocc$decile)

### create a numerical surge variable
mocc$surge <- 1*(mocc$decile=="0-9") +
  2*(mocc$decile=="10-19") +
  3*(mocc$decile=="20-29") +
  4*(mocc$decile=="30-39") +
  5*(mocc$decile=="40-49") +
  6*(mocc$decile=="50-59") +
  7*(mocc$decile=="60-69") +
  8*(mocc$decile=="70-79") +
  9*(mocc$decile=="80-89") +
  10*(mocc$decile=="90-100")

### center surge variable 
meansurge <- mean(mocc$surge, na.rm = TRUE)
minsurge <- min(mocc$surge, na.rm = TRUE)
maxsurge <- max(mocc$surge, na.rm = TRUE)
mocc$surge <- mocc$surge - mean(mocc$surge, na.rm = TRUE)
print(meansurge) 

### Four age categories using quartiles
age_quartiles <- quantile(mocc$age, probs=c(0.25, 0.50, 0.75),
                          na.rm=T)
age_quartiles
mocc$age_4cat <- 1*(mocc$age <= age_quartiles[1]) +
  2*(mocc$age > age_quartiles[1] & mocc$age <= age_quartiles[2]) +
  3*(mocc$age > age_quartiles[2] & mocc$age <= age_quartiles[3]) +
  4*(mocc$age > age_quartiles[3])

### exclude observations with missing values of relative dates
mocc <- mocc[!is.na(mocc$EventDate_MoccInit_Diff),]

### check if all observations were from MOCC states
### table(mocc$MOCC_Binary_Final)

### convert Date to Year and Month
mocc$Month_Yr <- format(as.Date(mocc$date), "%Y-%m")
### use [-3, 3] as week 0
mocc$week <- floor((mocc$EventDate_MoccInit_Diff+3)/7)

### days from 2019-01-01
mocc$days <- (as.numeric(as.POSIXct(mocc$date)-as.POSIXct(mocc$date[1]), units="days")) 
#mocc[mocc$days == 1459,]
#mocc[mocc$days == 1460,]
## last day was 2022-12-30

### determine the period in which data were available from all seven 
### Eight MOCC states with a MOCC initiation date on or after 6/1/2020 
### (Colorado, Maryland, Alaska, Utah, Virginia, Idaho, Oregon, NC)

mocc_initiation_dates <- c("2020-11-30", # Maryland
                          "2020-12-01", # Alaska
                          #"2020-04-16", # Arizona
                          "2020-11-01", # Colorado
                          #"2019-12-13", #Georgia
                          "2021-08-19", # Idaho
                          #"2020-05-28", # Minnesota
                          "2020-09-10", # North Carolina
                          "2021-08-21", # Oregon
                          "2021-08-01", # Utah
                          "2021-08-16") # Virginia 
                          #"2020-05-01" # Washington
mocc_initiation_dates <- as.Date(mocc_initiation_dates)

### indicator whether data were from Maryland
mocc$Maryland <- 1*(floor(as.numeric(as.POSIXct(mocc_initiation_dates[1]) -
             as.POSIXct(mocc$date), units="days")) == -mocc$EventDate_MoccInit_Diff)  
table(mocc$Maryland)

### indicator whether data were from Alaska
mocc$Alaska <- 1*(floor(as.numeric(as.POSIXct(mocc_initiation_dates[2]) -
             as.POSIXct(mocc$date), units="days")) == -mocc$EventDate_MoccInit_Diff)  
table(mocc$Alaska)

### indicator whether data were from Colorado

mocc$Colorado <- 1*(floor(as.numeric(as.POSIXct(mocc_initiation_dates[3]) -
             as.POSIXct(mocc$date), units="days")) == -mocc$EventDate_MoccInit_Diff)     
table(mocc$Colorado)

### indicator whether data were from Idaho

mocc$Idaho <- 1*(floor(as.numeric(as.POSIXct(mocc_initiation_dates[4]) -
             as.POSIXct(mocc$date), units="days")) == -mocc$EventDate_MoccInit_Diff)  
table(mocc$Idaho)

### indicator whether data were from North Carolina

mocc$NC <- 1*(floor(as.numeric(as.POSIXct(mocc_initiation_dates[5]) -
             as.POSIXct(mocc$date), units="days")) == -mocc$EventDate_MoccInit_Diff)  
table(mocc$NC)

### indicator whether data were from Oregon

mocc$Oregon <- 1*(floor(as.numeric(as.POSIXct(mocc_initiation_dates[6]) -
             as.POSIXct(mocc$date), units="days")) == -mocc$EventDate_MoccInit_Diff)  
table(mocc$Oregon)

### indicator whether data were from Utah

mocc$Utah <- 1*(floor(as.numeric(as.POSIXct(mocc_initiation_dates[7]) -
             as.POSIXct(mocc$date), units="days")) == -mocc$EventDate_MoccInit_Diff)  
table(mocc$Utah)

### indicator whether data were from Virginia

mocc$Virginia <- 1*(floor(as.numeric(as.POSIXct(mocc_initiation_dates[8]) -
             as.POSIXct(mocc$date), units="days")) == -mocc$EventDate_MoccInit_Diff)     
### Check the time period when data are available for each MOCC state
summary(mocc$week[mocc$Maryland==1])
summary(mocc$week[mocc$Alaska==1])
summary(mocc$week[mocc$Colorado==1])
summary(mocc$week[mocc$Idaho==1])
summary(mocc$week[mocc$NC==1])
summary(mocc$week[mocc$Oregon==1])
summary(mocc$week[mocc$Utah==1])
summary(mocc$week[mocc$Virginia==1])

### indicator whether data were from one of the eight MOCC states

mocc$eight_moccstates <- mocc$Maryland + mocc$Alaska + mocc$Colorado + 
  mocc$Idaho + mocc$Oregon + mocc$Utah + mocc$Virginia + mocc$NC

### indicator whether data were from one of the four MOCC states
mocc$four_moccstates <- mocc$Idaho + mocc$Oregon + mocc$Utah + mocc$Virginia 

### time period such that data from all eight MOCC states are available
minweek <- max(c(min(mocc$week[mocc$Maryland==1]),
    min(mocc$week[mocc$Alaska==1]),
    min(mocc$week[mocc$Colorado==1]),
    min(mocc$week[mocc$Idaho==1]),
    min(mocc$week[mocc$NC==1]),
    min(mocc$week[mocc$Oregon==1]),
    min(mocc$week[mocc$Utah==1]),
    min(mocc$week[mocc$Virginia==1])))

maxweek <- min(c(max(mocc$week[mocc$Maryland==1]),
    max(mocc$week[mocc$Alaska==1]),
    max(mocc$week[mocc$Colorado==1]),
    max(mocc$week[mocc$Idaho==1]),
    max(mocc$week[mocc$NC==1]),
    max(mocc$week[mocc$Oregon==1]),
    max(mocc$week[mocc$Utah==1]),
    max(mocc$week[mocc$Virginia==1])))

mindiff_pre <- min(as.numeric(as.POSIXct(mocc_initiation_dates)-
             as.POSIXct(as.Date("2019-01-01")), units="days"))

summary(as.numeric(as.POSIXct(mocc_initiation_dates)-
             as.POSIXct(as.Date("2022-12-30")), units="days"))
mindiff_post <- min(-as.numeric(as.POSIXct(mocc_initiation_dates)-
             as.POSIXct(as.Date("2022-12-30")), units="days"))

### Create a data set with data from the eight MOCC states only that were
### available at each (relative) week
mocc_v2 <- mocc[mocc$eight_moccstates==1,]
mocc_v2 <- mocc_v2[mocc_v2$week >= minweek
                 & mocc_v2$week <= maxweek,]
### check if the first week and the last week were full week
summary(mocc_v2$week)
min_week <- min(mocc_v2$week)
max_week <- max(mocc_v2$week)

### full week in week -14
table(mocc$date[mocc$NC == 1 & mocc$week == -14])

table(mocc$date[mocc$Idaho == 1 & mocc$week == 71])

### Remove data from week -14 and 71 since they are not full weeks
mocc_v2 <- mocc_v2[mocc_v2$week > min_week,]
mocc_v2 <- mocc_v2[mocc_v2$week < max_week,]

summary(mocc_v2$week)
### table(mocc_v2$week)

### check states and weeks with missing surge data
table(is.na(mocc_v2$surge[mocc_v2$Maryland==1]), mocc_v2$week[mocc_v2$Maryland==1])
table(is.na(mocc_v2$surge[mocc_v2$Alaska==1]), mocc_v2$week[mocc_v2$Alaska==1])
table(is.na(mocc_v2$surge[mocc_v2$Colorado==1]), mocc_v2$week[mocc_v2$Colorado==1])

table(is.na(mocc_v2$surge[mocc_v2$Idaho==1]), mocc_v2$week[mocc_v2$Idaho==1])
table(is.na(mocc_v2$surge[mocc_v2$NC==1]), mocc_v2$week[mocc_v2$NC==1])
table(is.na(mocc_v2$surge[mocc_v2$Oregon==1]), mocc_v2$week[mocc_v2$Oregon==1])
table(is.na(mocc_v2$surge[mocc_v2$Utah==1]), mocc_v2$week[mocc_v2$Utah==1])
table(is.na(mocc_v2$surge[mocc_v2$Virginia==1]), mocc_v2$week[mocc_v2$Virginia==1])

### Combine Rural   Suburban   Wilderness
mocc_v2$Urbanicity_2cat <- 1*(mocc_v2$urbanicity=="Urban") +
  2*(mocc_v2$urbanicity != "Urban")
mocc_v2$Urbanicity_2cat[is.na(mocc_v2$urbanicity)] <- NA

mocc_v2$Urbanicity_2cat <-  factor(mocc_v2$Urbanicity_2cat,
       labels = c("Urban", "Non-Urban"))
### All eight states initiated MOCC post 1st wave
### no need to conduct separate analysis for Pandemic period and Post 1st Wave period

dim(mocc_v2)
### Total number of IFTs using data from seven MOCC states
table(mocc_v2$week)

### save the data to a csv file
write.csv(mocc_v2, file="mocc_eightstates_final.csv")

```


\bf ITS Analysis: full data from eight MOCC states

```{r, echo=TRUE}
source("MOCC_Routines_final.R")
library(ggplot2)
fitnb_full <- mocc_routines_v4(mocc_v2, "full_1.pdf", 
              "full_2.pdf", model_selected = NULL)
library(knitr)
kable(round(fitnb_full$fit_nb_coefficients, digits=4), caption="Interrupted time series analysis (logarithm of rate ratio) under the negative-binomial model")

kable(round(fitnb_full$fit_nb_exp_coefficients, digits=4), caption="Interrupted time series analysis (rate ratio) under the negative-binomial model")

kable(round(fitnb_full$parest, digits=4), caption="Effects of surge and MOCC initiation on the logarithm of IFT rate")

kable(round(fitnb_full$parest_exp, digits=4), caption="Effects of surge and MOCC initiation on the IFT rate")

### Creating scatter plot with its confidence intervals 
gfg<-round(data.frame(x = c(-1:7)*10, 
                      y = fitnb_full$parest_exp[1:9,1], 
                      low = fitnb_full$parest_exp[1:9,2], 
                      up = fitnb_full$parest_exp[1:9,3]), 4) 
ggplot(gfg, aes(x = factor(x), y = y )) + geom_point() + 
  geom_errorbar(aes(ymin = low, 
                    ymax = up)) + 
  scale_y_continuous(trans='log') +
  xlab("Week") +
  ylab("Surge Effect on IFT Rate") 
  ggsave("surge_effect_full.pdf")

  gfg<-round(data.frame(x = 1:10, 
                      y = fitnb_full$parest_exp[10:19,1], 
                      low = fitnb_full$parest_exp[10:19,2], 
                      up = fitnb_full$parest_exp[10:19,3]), 4) 
ggplot(gfg, aes(x = factor(x), y = y )) + geom_point() + 
  geom_errorbar(aes(ymin = low, 
                    ymax = up)) + 
  scale_y_continuous(breaks=c(0, 1, 5, 10, 15, 20, 25), trans='log') +
  xlab("Surge") +
  ylab("Immediate MOCC Effect on IFT Rate") 
  ggsave("immediate_MOCC_effect_full.pdf")

    gfg<-round(data.frame(x = 1:10, 
                      y = fitnb_full$parest_exp[20:29,1], 
                      low = fitnb_full$parest_exp[20:29,2], 
                      up = fitnb_full$parest_exp[20:29,3]), 4) 
ggplot(gfg, aes(x = factor(x), y = y )) + geom_point() + 
  geom_errorbar(aes(ymin = low, 
                    ymax = up)) + 
  scale_y_continuous(trans='log') +
  xlab("Surge") +
  ylab("Sustained MOCC Effect on IFT Rate") 
  ggsave("sustained_MOCC_effect_full.pdf")


### predicted IFT rate (and 95 CIs) from relative 
### weeks -13 to 70 at specified surge decile and season   
library(xlsx)

for (surge in 1:10)
  for (season in c("Spring", "Summer", "Fall", "Winter"))
  {
    preds <- mocc_pred(fit_nb=fitnb_full$fit_nb, covmat=fitnb_full$covmat,
          data_ad=fitnb_full$data_ad, 
          surge=surge, meansurge=meansurge, season=season)
    if (surge == 1 & season == "Spring")
    {
      write.xlsx(data.frame(preds), file="MOCC_pred_full.xlsx",
                 sheetName = paste("surge_", surge, "_", season, sep=""),
                 row.names = FALSE) 
      } else
      {
        write.xlsx(data.frame(preds), file="MOCC_pred_full.xlsx",
                 sheetName = paste("surge_", surge, "_", season, sep=""),
                 row.names = FALSE, append=TRUE) 
      }
  }
  

```


\bf ITS Analysis: data from each category of Urbancity: Rural, Suburban, Urban, Wilderness

```{r, echo=TRUE}
taburbancity <- table(mocc_v2$urbanicity)
rownames_urbancity <- rownames(taburbancity)
for (i in 1:length(taburbancity))
{
  print(rownames_urbancity[i])
  print(mocc_routines_v4(mocc_v2[mocc_v2$urbanicity == rownames_urbancity[i],], paste(rownames_urbancity[i],"_1.pdf",sep=""),
                paste(rownames_urbancity[i],"_2.pdf",sep=""), model_selected = 3))
}
```

\bf ITS Analysis: data from each category of Urbancity\_2cat:  Urban vs Non-Urban (Rural, Suburban, Wilderness)

```{r, echo=TRUE}
taburbancity_2cat <- table(mocc_v2$Urbanicity_2cat)
rownames_urbancity_2cat <- rownames(taburbancity_2cat)
for (i in 1:length(taburbancity_2cat))
{
  print(rownames_urbancity_2cat[i])
  print(mocc_routines_v4(mocc_v2[mocc_v2$Urbanicity_2cat == rownames_urbancity_2cat[i],], paste(rownames_urbancity_2cat[i],"_1.pdf",sep=""),
                paste(rownames_urbancity_2cat[i],"_2.pdf",sep=""), model_selected = 3))
}
```

\bf ITS Analysis: data from each category of Region: Midwest, South, West (algorithm doesn't converge)

```{r, echo=TRUE}
tabregion <- table(mocc_v2$region)
rownames_region <- rownames(tabregion)
for (i in 1:length(tabregion))
{
  print(rownames_region[i])
  print(mocc_routines_v4(mocc_v2[mocc_v2$region == rownames_region[i],], paste(rownames_region[i],"_1.pdf",sep=""),
                paste(rownames_region[i],"_2.pdf",sep=""), model_selected = 3))
}
```

\bf ITS Analysis: data from each category of age groups: <=q1, (q1, q2], (q2, q3], >q3

```{r, echo=TRUE}
tabage <- table(mocc_v2$age_4cat)
rownames_age <- rownames(tabage)
for (i in 1:length(tabage))
{
  print(rownames_age[i])
  print(mocc_routines_v4(mocc_v2[mocc_v2$age_4cat == rownames_age[i],], paste("age",rownames_age[i],"_1.pdf",sep=""),
                paste("age",rownames_age[i],"_2.pdf",sep=""), model_selected = 3))
}
```

\bf ITS Analysis: data from each category of transport mode: air and ground

```{r, echo=TRUE}
tmptab <- table(mocc_v2$transport_mode)
tmprownames <- rownames(tmptab)
for (i in 1:length(tmptab))
{
  print(tmprownames[i])
  print(mocc_routines_v4(mocc_v2[mocc_v2$transport_mode == tmprownames[i],], paste(tmprownames[i],"_1.pdf",sep=""),
                paste(tmprownames[i],"_2.pdf",sep=""), model_selected = 3))
}
```

\bf ITS Analysis: data from each category of level of care: ALS, BLS, specialty

```{r, echo=TRUE}
tmptab <- table(mocc_v2$level_of_care)
tmprownames <- rownames(tmptab)
for (i in 1:length(tmptab))
{
  print(tmprownames[i])
  print(mocc_routines_v4(mocc_v2[mocc_v2$level_of_care == tmprownames[i],], paste(tmprownames[i],"_1.pdf",sep=""),
                paste(tmprownames[i],"_2.pdf",sep=""), model_selected = 3))
}
```

\bf ITS Analysis: data from each category of inital patient acuity: black, green, red, unknown, yellow

```{r, echo=TRUE}
table(mocc$initial_acuity)
### too few black, combined with unknown
mocc_v2$initial_acuity[mocc_v2$initial_acuity=="black"] <- 
  "unknown"
tmptab <- table(mocc_v2$initial_acuity)
tmprownames <- rownames(tmptab)
for (i in 1:length(tmptab))
{
  print(tmprownames[i])
  print(mocc_routines_v4(mocc_v2[mocc_v2$initial_acuity == tmprownames[i],], paste(tmprownames[i],"_1.pdf",sep=""),
                paste(tmprownames[i],"_2.pdf",sep=""), model_selected = 3))
}
```